Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit

### 2024-06-28 optim.tex
half-way restructuing section 4 and 5, add illustration

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{algorithm}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal of Multivariate Analysis}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num-names}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit},
  pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Wai Yin Leung},
  pdfkeywords={projection pursuit, jellyfish search optimiser
(JSO), optimisation, grand tour, high-dimensional data, exploratory data
analysis},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{keyword}
    projection pursuit \sep jellyfish search optimiser
(JSO) \sep optimisation \sep grand tour \sep high-dimensional data \sep 
    exploratory data analysis
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt, frame hidden, sharp corners, enhanced]}{\end{tcolorbox}}\fi

@@

\hypertarget{sec-background}{%
\section{Projection pursuit, tours, index functions and optimisation
{[}Di and Sherry{]}}\label{sec-background}}

@@

\begin{tcolorbox}[enhanced jigsaw, colback=white, opacitybacktitle=0.6, left=2mm, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black, titlerule=0mm, bottomtitle=1mm, bottomrule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, opacityback=0, arc=.35mm, breakable, rightrule=.15mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm]

@@

\begin{equation}\protect\hypertarget{eq-proj-dist}{}{d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}\ .}\label{eq-proj-dist}\end{equation}

@@

\begin{equation}\protect\hypertarget{eq-squintability}{}{
\varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)\ .
}\label{eq-squintability}\end{equation}

@@

\begin{equation}\protect\hypertarget{eq-squintability-parametric}{}{
\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
}\label{eq-squintability-parametric}\end{equation}

@@

\hypertarget{sec-vis}{%
\section{Visualisation of jellyfish optimiser}\label{sec-vis}}

@@

The JSO hyper-parameters used are 50 jellyfishes (\texttt{n\_jellies})
and a maximum of 50 tries (\texttt{max\_tries}). The full data contains
50 simulations, indexed by the variable \texttt{sim}, with the
associated \texttt{seed}s. Recorded in columns \texttt{basis},
\texttt{index\_val}, \texttt{tries}, \texttt{loop} and \texttt{time},
are the sampled projection basis, index value, iteration ID, jellyfish
ID, and computational time respectively. The basis column records every
visited basis before comparing with the current (last) basis. If a
visited basis has a smaller index value than the current (last) one, JSO
will keep the current basis for the next iteration in the search but
still records the visited inferior basis. This data can be used to
derive numerical summaries and plots:

\begin{itemize}
\item
  \textbf{Success rate}: The success rate is defined as the proportion
  of simulations that achieves a final index value within 0.05 of the
  best index value found among all 50 simulations. In the dataset above,
  the optimal index value is 0.999. Out of the 50 simulations, 44
  achieved an index value within {[}0.949, 0.999{]}, resulting in a
  success rate of 0.88. Figure~\ref{fig-projection} shows the best
  projection found in each simulation, sorted by the index value. Most
  of the projections show a clear sine-wave structure, while the last
  six show only a straight line with clusters at the end. When local
  optima exist in the projection pursuit problem, plotting the
  projections found in each simulation or by each jellyfish can help
  identify these local optima.
\item
  \textbf{Projection distance}: The projection distance between
  projection bases and the optimal best basis (or the empirical best
  basis if the optimal is unknown) measures how close a visited basis is
  to the best basis and it is calculated as the Frobenius norm in
  Equation

  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \tightlist
  \item
    Plotting projection distances against index values for the bases
    visited reveals how the index value changes as the basis approaches
    the optimal basis, helping to calculate the squintability measure.
    Figure~\ref{fig-index-value-proj-dist} shows the index value against
    the projection distance for the given dataset and its binned
    version. As the projection distance decreases, the index value
    initially increases with an increasing slope and then levels off.
    The index value shows greater variation at medium projection
    distances compared to small and large distances.
  \end{enumerate}
\end{itemize}

@@

{\centering \includegraphics{optim_files/figure-pdf/fig-projection-1.pdf}

@@

\caption{\label{fig-projection}Best projections found by the JSO in each
of the 50 simulations, sorted by index values. The projection pursuit
problem is to detect the 6D sine-wave structure using the spline index.
Most simulations have detected a clear sine-wave pattern, while the last
six show a straight line with clustering at the ends. In this
simulation, the success rate is 0.88 (44/50).}

\end{figure}

@@

{\centering \includegraphics{optim_files/figure-pdf/fig-index-value-proj-dist-1.pdf}

@@

\caption{\label{fig-index-value-proj-dist}Projection distance plotted
against the index value for detecting the 6D sine-wave structure using
the spline index function. a) The index values start around 0 with low
variance at large projection distances. As the projection distance
decreases, the index values increase with a larger variance. Towards the
minimum projection distance, the index values increase and cluster
around 1, indicating the optimiser is approaching to the optimum. b)
Projection distances are binned with a bin width of 0.1, and index
values are averaged over each bin. The relationship between the two
variables forms a sigmoid curve.}

\end{figure}

\hypertarget{sec-app}{%
\section{Application {[}Di and Sherry{]}}\label{sec-app}}

The JSO provides a new swarm-based search strategy to optimise
projection pursuit problems. Two examples are provided in this section
to demonstrate its performance. The first example compares the
performance of JSO with an exist optimiser, search better, and explores
how JSO behaves with different combinations of hyper-parameters. The
second example studies the effect of optimisation properties defined in
Section~\ref{sec-theory}, along with JSO hyper-parameters, on the
outcome (success rate) of the optimisation.

\hypertarget{sec-app-1}{%
\subsection{Performance of JSO in pipe-finding
problems}\label{sec-app-1}}

The performance of JSO is investigated in two folds: 1) compare it with
a commonly used optimiser: the search-better optimiser (SBO)
\citep{RJ-2021-105, laa_using_2020}, and 2) examine its success rate
under different hyper-parameters (number of jellyfishes and maximum
number of tries). The projection pursuit problem used is finding the
pipe shape using the holes index, investigated by
\citet{laa_using_2020}.

@@

\end{figure}

Fifty simulations are conducted with both JSO and the SBO, in four
dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100 jellyfishes and 100
maximum number of tries, while the SBO allows a maximum of 1000 samples
at each iteration before the algorithm terminates. Figure~\ref{fig-proj}
presents the final projections found by the two optimisers, broken down
by 10th quantile, faceted by the data dimension. In the 6-dimensional
data scenario, JSO consistently identifies a clear pipe shape. The SBO
also finds the pipe shape but with a wide rim, suggesting a further
polish search may be required. With increasing dimensions, JSO may not
always identify the pipe shape due to random sampling, but it still
presents a pipe shape in over 50\% of cases. When compared to the SBO
JSO reaches higher index values and clearer pipe shapes across all
quantiles.

In the second experiment, fifty simulations are conducted at each
hyper-parameter combination to analyse its effects on the JSO success
rate. The hyper-parameters tested include: 1) 20, 50, and 100
jellyfishes, and 2) 50 and 100 maximum number of tries. In each
simulation, the success rate is calculated as defined in
Section~\ref{sec-vis}. Figure~\ref{fig-proportion} presents the
proportion of success for each hyper-parameter combinations. As the
number of jellyfishes and maximum tries increase, the success rate also
increases. For simpler problems (6 dimensions), small parameter values
(20 jellyfishes and a maximum number of tries of 50) can already result
in high success rate, while for higher-dimensional problems (8, 10, and
12 dimensions), a combination of 100 jellyfishes and 100 maximum number
of tries is necessary for at least half of the jellyfishes to find the
projection close to the optimal.

\hypertarget{sec-app-2}{%
\subsection{Factors affecting JSO success rate: optimisation properties
and jellyfish hyper-parameters}\label{sec-app-2}}

The optimisation properties, including smoothness, squintability, and
speed, offer numerical metrics to characterise the complexity of
projection pursuit optimisation problems. This example investigates how
these metrics, along with the hyper-parameters affect the JSO success
rate. Simulations are conducted to obtain the success rate across a
wider array of projection pursuit problems, characterised by
shape-to-find, data dimension, and the index function, as well as
different hyper-parameter combinations. Smoothness and squintability are
calculated for each projection pursuit problem as described in
Section~\ref{sec-theory}. A generalised linear model is used to
construct the relationship between the success rate, JSO
hyper-parameters and optimisation properties.

In addition to the pipe-finding problem, a new shape, sine wave, is
investigated in 6D and 8D spaces with six indexes considered:
\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
\texttt{spline}, and \texttt{stringy}. Combining with two JSO
hyper-parameters, a total of 52 cases is produced, comprising of 24
pipe-finding cases and 28 sine-wave finding cases. For each case, JSO is
run fifty times and the success rate is calculated as in
Section~\ref{sec-app-1}.

Smoothness and squintability are computed for each case, following the
procedures outlined in Section~\ref{sec-smoothness} and
Section~\ref{sec-squintability}. To calculate smoothness, three hundred
random bases are simulated. Index values and projection distance (to the
optimal basis) are calculated for each random basis before fitting into
a Gaussian process model to obtain the smoothness measure. For
squintability, fifty random bases are sampled and interpolated to the
optimal basis with a step size of 0.005. Index values and projection
distances are calculated for these interpolated bases and the index
values are averaged with a bin width of 0.005. The squintability measure
is then calculated from fitting a 4-parameter scaled logistic function
of index value against projection distance, estimated by non-linear
least squares.

@@

\end{figure}

\hypertarget{tbl-smoothness-squintability}{}
\begin{table}
\caption{\label{tbl-smoothness-squintability}Parameters estimated from the Gaussian process (including variance,
range, smoothness, and nugget) and scaled logistic function
(\(\theta_1\) to \(\theta_4\)) for the pipe-finding and sine-wave
finding problems. The column ``smooth'' and ``squint'' represent the
smoothness and squintability measures. }\tabularnewline

@@

\hypertarget{tbl-mod-output}{}
\begin{longtable}[t]{|>{}lrrr>{}r|}
\caption{\label{tbl-mod-output}Model estimates of proportion of jellyfish success on optimisation
properties and jellyfish hyper-parameters from the generalised linear
model with a binomial family and a logit link function. The variable
smoothness, squintability, number of jellyfish and maximum number of
tries are positively associated with JSO success rate while data
dimension and being flagged as long runtime are negatively associated
with the success rate. The variable squintability and dimension are
significant, suggesting their importance relative to jellyfish
hyper-parameters in the optimisation success. }\tabularnewline

\toprule
term & estimate & std.error & statistic & p.value\\
\midrule
Intercept & -4.523 & 5.332 & -0.848 & 0.396\\
Smoothness & 1.195 & 1.915 & 0.624 & 0.533\\
Squintability & 2.060 & 0.684 & 3.011 & 0.003\\
dimension (d) & -0.628 & 0.255 & -2.462 & 0.014\\
long time & -0.874 & 1.286 & -0.680 & 0.497\\
\addlinespace
number of jellyfish & 0.216 & 0.127 & 1.701 & 0.089\\
maximum number of tries & 0.113 & 0.150 & 0.752 & 0.452\\
\bottomrule
\end{longtable}
\endgroup{}

@@

\hypertarget{sec-conclusion}{%
\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}}

>>>>

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{algorithm}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Journal of Multivariate Analysis}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num-names}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit},
  pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Wai Yin Leung},
  pdfkeywords={projection pursuit, jellyfish search optimiser
(JSO), optimisation, grand tour, high-dimensional data, exploratory data
analysis},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{keyword}
    projection pursuit \sep jellyfish search optimiser
(JSO) \sep optimisation \sep grand tour \sep high-dimensional data \sep 
    exploratory data analysis
\end{keyword}
\end{frontmatter}

@@

\section{Projection pursuit, tours, index functions and optimisation
{[}Di and Sherry{]}}\label{sec-background}

@@

\begin{tcolorbox}[enhanced jigsaw, left=2mm, coltitle=black, bottomtitle=1mm, arc=.35mm, colbacktitle=quarto-callout-note-color!10!white, colback=white, bottomrule=.15mm, opacitybacktitle=0.6, breakable, opacityback=0, titlerule=0mm, rightrule=.15mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, toprule=.15mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm]

@@

\begin{equation}\phantomsection\label{eq-proj-dist}{d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}\ .}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-squintability}{
\varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)\ .
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-squintability-parametric}{
\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
}\end{equation}

@@

\section{Visualisation of jellyfish optimiser}\label{sec-vis}

The JSO provides a new swarm-based search strategy to optimise
projection pursuit problems. This section outlines two simulation setups
to demonstrate the performance of JSO. In the first example, the
performance of JSO is compared with an exist optimiser, Creeping Random
Search (CRS), and explores how JSO behaves with different combinations
of hyper-parameters. The second example studies the effect of
optimisation properties defined in Section~\ref{sec-theory}, along with
JSO hyper-parameters, on the outcome (success rate) of the optimisation.

@@

The JSO hyper-parameters used are 50 jellyfishes (\texttt{n\_jellies})
and a maximum of 50 tries (\texttt{max\_tries}). The full data contains
50 simulations, indexed by the variable \texttt{sim}, with the
associated \texttt{seed}s. Recorded in columns \texttt{basis},
\texttt{index\_val}, \texttt{tries}, \texttt{loop} and \texttt{time},
are the sampled projection basis, index value, iteration ID, jellyfish
ID, and computational time respectively. The basis column records every
visited basis before comparing with the current (last) basis. If a
visited basis has a smaller index value than the current (last) one, JSO
will keep the current basis for the next iteration in the search but
still records the visited inferior basis.

\subsection{Performance of JSO in pipe-finding
problems}\label{sec-app-1}

The performance of JSO is investigated in two folds: 1) compare it with
a commonly used optimiser: the search-better optimiser (SBO)
\citep{RJ-2021-105, laa_using_2020}, and 2) examine its success rate
under different hyper-parameters (number of jellyfishes and maximum
number of tries). The projection pursuit problem used is finding the
pipe shape using the holes index, investigated by
\citet{laa_using_2020}. Fifty simulations are conducted with both JSO
and the SBO, in four dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100
jellyfishes and 100 maximum number of tries, while the SBO allows a
maximum of 1000 samples at each iteration before the algorithm
terminates.

In the second experiment, fifty simulations are conducted at each
hyper-parameter combination to analyse its effects on the JSO success
rate. The hyper-parameters tested include: 1) 20, 50, and 100
jellyfishes, and 2) 50 and 100 maximum number of tries. The success rate
is defined as the proportion of simulations that achieves a final index
value within 0.05 of the best index value found among all 50
simulations. In the dataset above, the optimal index value is 0.999. Out
of the 50 simulations, 44 achieved an index value within {[}0.949,
0.999{]}, resulting in a success rate of 0.88.

\subsection{Factors affecting JSO success rate: optimisation properties
and jellyfish hyper-parameters}\label{sec-app-2}

In addition to the pipe-finding problem, a new shape, sine wave, is
investigated in 6D and 8D spaces with six indexes considered:
\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
\texttt{spline}, and \texttt{stringy}. Combining with two JSO
hyper-parameters, a total of 52 cases is produced, comprising of 24
pipe-finding cases and 28 sine-wave finding cases. For each case, JSO is
run fifty times and the success rate is calculated.

Smoothness and squintability are computed for each case, following the
procedures outlined in Section~\ref{sec-smoothness} and
Section~\ref{sec-squintability} and illustrated in
Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}. To
calculate smoothness, three hundred random bases are simulated. Index
values and projection distance (to the optimal basis) are calculated for
each random basis before fitting into a Gaussian process model to obtain
the smoothness measure. For squintability, fifty random bases are
sampled and interpolated to the optimal basis with a step size of 0.005.
Index values and projection distances are calculated for these
interpolated bases and the index values are averaged with a bin width of
0.005. The squintability measure is then calculated from fitting a
4-parameter scaled logistic function of index value against projection
distance, estimated by non-linear least squares.

@@

\centering{

\includegraphics[width=6.77in,height=\textheight]{figures/smoothness.png}

@@

\caption{\label{fig-smoothness}Illustration of steps to calculate
smoothness.}

\end{figure}%

@@

\centering{

\includegraphics[width=6.77in,height=\textheight]{figures/squintability.png}

@@

\caption{\label{fig-squintability}Illustration of steps to calculate
squintability. For all the bases, calculate the projection distance to
the optimal basis and the corresponding index value. Bin the index value
by the projection distance, then claculate the average index value for
each bin. Fit the scaled sigmoid function from Equation
\ref{eq-squintability} to the binned index values against the projection
distance using non-linear least square. Use Equation \ref{eq-parametric}
to calculate the squintability measure from the estimated \(\theta_1\)
to \(\theta_4\).}

\end{figure}%

\section{Results {[}Di and Sherry{]}}\label{sec-app}

\subsection{Performance of JSO in pipe-finding
problems}\label{sec-app-1}

Figure~\ref{fig-proj} presents the final projections found by the two
optimisers, broken down by 10th quantile, faceted by the data dimension.
In the 6-dimensional data scenario, JSO consistently identifies a clear
pipe shape. The SBO also finds the pipe shape but with a wide rim,
suggesting a further polish search may be required. With increasing
dimensions, JSO may not always identify the pipe shape due to random
sampling, but it still presents a pipe shape in over 50\% of cases. When
compared to the SBO JSO reaches higher index values and clearer pipe
shapes across all quantiles.

Figure~\ref{fig-proportion} presents the proportion of success for each
hyper-parameter combinations. As the number of jellyfishes and maximum
tries increase, the success rate also increases. For simpler problems (6
dimensions), small parameter values (20 jellyfishes and a maximum number
of tries of 50) can already result in high success rate, while for
higher-dimensional problems (8, 10, and 12 dimensions), a combination of
100 jellyfishes and 100 maximum number of tries is necessary for at
least half of the jellyfishes to find the projection close to the
optimal.

@@

\end{figure}%

\subsection{Factors affecting JSO success rate: optimisation properties
and jellyfish hyper-parameters}\label{sec-app-2}

@@

\end{figure}%

\begin{table}

\caption{\label{tbl-smoothness-squintability}Parameters estimated from
the Gaussian process (including variance, range, smoothness, and nugget)
and scaled logistic function (\(\theta_1\) to \(\theta_4\)) for the
pipe-finding and sine-wave finding problems. The column ``smooth'' and
``squint'' represent the smoothness and squintability measures.}

\centering{

@@

\begin{longtable}[t]{|>{}lrrr>{}r|}

\caption{\label{tbl-mod-output}Model estimates of proportion of
jellyfish success on optimisation properties and jellyfish
hyper-parameters from the generalised linear model with a binomial
family and a logit link function. The variable smoothness,
squintability, number of jellyfish and maximum number of tries are
positively associated with JSO success rate while data dimension and
being flagged as long runtime are negatively associated with the success
rate. The variable squintability and dimension are significant,
suggesting their importance relative to jellyfish hyper-parameters in
the optimisation success.}

\tabularnewline

\toprule
term & estimate & std.error & statistic & p.value\\
\midrule
Intercept & -4.523 & 5.332 & -0.848 & 0.396\\
Smoothness & 1.195 & 1.915 & 0.624 & 0.533\\
Squintability & 2.060 & 0.684 & 3.011 & 0.003\\
dimension (d) & -0.628 & 0.255 & -2.462 & 0.014\\
long time & -0.874 & 1.286 & -0.680 & 0.497\\
\addlinespace
number of jellyfish & 0.216 & 0.127 & 1.701 & 0.089\\
maximum number of tries & 0.113 & 0.150 & 0.752 & 0.452\\
\bottomrule

\end{longtable}

\endgroup{}

@@

\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}
